Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-05-20 paper.tex
Add Yonglin's notes and citations from email Apr 26

%\subsubsection*{Why do scattering and pair processes matter?}
Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
modify neutrino distributions in angle and dilute the emitted spectrum over a
larger emitting surface \citep{pere2016-asl}.
This is especially pronounced in the case of heavy-lepton neutrinos.
Inelastic scattering and pair processes can introduce further modifications.
Any phenomena that involve neutrino-neutrino interactions,
for example neutrino oscillations
\cite{duan2010-collective,malk2012-mnr_1,vlas2018-multiangle}
and neutrino-antineutrino annihilation \cite{asan2000-nunubar},
depend sensitively on the angular distribution.
And spectral changes in neutrino distributions can strongly affect the nuclear
processes occuring in the ejected and irradiated material
\cite{fouc2016-m1_evolve_n}.
The ray tracing method we present in this paper captures the dominant effects
of elastic scattering, while leaving the physics of inelastic scattering and
pair processes for later work.

@@

However, we can solve a similar but tractable problem along a single ray.
Assume all the neutrino rays intersecting an event along a given test ray
have undergone the same flavor evolution history as that of the test ray,
i.e.\ $S$ is the same for all rays sharing that event.
This is the so-called single-angle approximation.
\footnote{TODO (Gail, Yonglin): citation?}
Though widely used in the supernova oscillation literature
and shown to be qualitatively faithful in those environments,
\footnote{TODO (Gail, Yonglin): citation?}
we note that a recent study has shown that the single-angle approximation
may mask certain instabilities in the flavor evolution
\cite{wu2017-fast_neutrino_conversions}.

@@

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
In Fig.~\ref{fig:survival-scat}
we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the \emph{scat} treatment. (There is no flavor transformation
in the \emph{noscat} treatment.)
The survival probabilities are the terms of
the flavor evolution matrix $S$ that describe how much of the original flux
of a given flavor survives a transition.
\footnote{TODO (Yonglin): define $P_{\nu_\alpha\rightarrow\nu_\alpha}$
  in terms of $S$}
Where $P_{\nu_e\rightarrow\nu_e}$ decreases, for example in
Fig.~\ref{fig:survival-scat} outside of $r\sim400$~km,
a fraction of the original $\nu_e$ content of the beam oscillates into
$\nu_\mu$ and $\nu_\tau$, and a fraction of $\nu_\mu$ and $\nu_\tau$
oscillates into $\nu_e$.

>>>>

%\subsubsection*{Why do scattering and pair processes matter?}
Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
modify neutrino distributions in angle and dilute the emitted spectrum over a
larger emitting surface \citep{pere2016-asl}.
This is especially pronounced in the case of heavy-lepton neutrinos.
Inelastic scattering and pair processes can introduce further modifications.
Any phenomena that involve neutrino-neutrino interactions,
for example neutrino oscillations
\cite{duan2010-collective,malk2012-mnr_1,vlas2018-multiangle}
and neutrino-antineutrino annihilation \cite{asan2000-nunubar},
depend sensitively on the angular distribution.
And spectral changes in neutrino distributions can strongly affect the nuclear
processes occuring in the ejected and irradiated material
\cite{surm2006-grb_nucsynth,malk2012-mnr_1,caba2012-nu_spectra,fouc2016-m1_evolve_n}.
The ray tracing method we present in this paper captures the dominant effects
of elastic scattering, while leaving the physics of inelastic scattering and
pair processes for later work.

@@

However, we can solve a similar but tractable problem along a single ray.
We assume that all the neutrino rays intersecting an event along a given test ray
have undergone the same flavor evolution history as that of the test ray,
i.e.\ $S$ is the same for all rays sharing that event.
This is the so-called single-angle approximation.
Though widely used in the supernova oscillation literature
and shown to be qualitatively faithful in those environments,
we note that recent studies have discovered configurations in which
the single-angle approximation masks certain instabilities
in the flavor evolution \cite{wu2017-fast_neutrino_conversions},
or introduces qualitatively different flavor evolution behavior
than that of the full multi-angle calculation \cite{vlas2018-multiangle}.

@@

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
In Fig.~\ref{fig:survival-scat}
we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the \emph{scat} treatment. (There is no flavor transformation
in the \emph{noscat} treatment. \footnote{TODO (Brett): ubpdate with new plots})
The survival probability is the probability that a neutrino, if measured,
will be found to be in its original flavor state.
The survival probabilities are computed at each point along the trajectory
from the flavor evolution matrix,
$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}(E)|^2$.
When $P_{\nu_e \to \nu_e}$ decreases,
as can be seen for example in Fig.~\ref{fig:survival-scat} for $s>400$~km,
some of the electron-type neutrinos have oscillated into
mu- or tau-type neutrinos.
